### 2017-04-04 sec-conclusion.tex
Switch to 2L-HMM for the model abbreviation.
Update the reviewer response draft to mention this concrete change.

Currently, the model learns a transition matrix over the latent states that
is \emph{shared} across all students. It would be interesting to instead
learn a different latent state transition matrix for each individual
student, but keep the second-level Markov models shared. This would provide
the model with more flexibility which may be desirable itself, but would
also naturally result in a description of a student (via his or her HMM
transitions) that could be incorporated into existing supervised learning
techniques that try to predict student outcomes for understanding
which of the latent behavior patterns discovered by TL-HMM are most predictive
of the performance of student learning.

>>>>

Currently, the model learns a transition matrix over the latent states that
is \emph{shared} across all students. It would be interesting to instead
learn a different latent state transition matrix for each individual
student, but keep the second-level Markov models shared. This would provide
the model with more flexibility which may be desirable itself, but would
also naturally result in a description of a student (via his or her HMM
transitions) that could be incorporated into existing supervised learning
techniques that try to predict student outcomes for understanding
which of the latent behavior patterns discovered by 2L-HMM are most predictive
of the performance of student learning.

### 2017-04-04 sec-intro.tex
Switch to 2L-HMM for the model abbreviation.
Update the reviewer response draft to mention this concrete change.

Thus, what we propose is a \emph{latent variable approach} to mining student behavior
patterns that is \emph{probabilistic} for inference and 
\emph{flexible to model state changes over different time resolutions}. 
More specifically, we propose a novel two-layer hidden Markov model (TL-HMM) to
 discover latent student behavior patterns via
unsupervised learning on large collections of student behavior observation
sequences. Evaluation results on a MOOC data set on Coursera demonstrate that TL-HMM can effectively 
discover a variety of interesting interpretable student behavior patterns at different levels of resolution, many of which are beyond what existing approaches can discover. TL-HMM further enables easy use of the discovered patterns for discriminative analysis such as prediction of learning outcome. 
Since our proposed methods are unsupervised, they can potentially be applied to any MOOC data without requiring any manual work to facilitate understanding of student behaviors and their variations, opening up many possibilities for developing intelligent tutoring systems that can adaptive to student behavior.

>>>>

Thus, what we propose is a \emph{latent variable approach} to mining student behavior
patterns that is \emph{probabilistic} for inference and 
\emph{flexible to model state changes over different time resolutions}. 
More specifically, we propose a novel two-layer hidden Markov model (2L-HMM) to
 discover latent student behavior patterns via
unsupervised learning on large collections of student behavior observation
sequences. Evaluation results on a MOOC data set on Coursera demonstrate that 2L-HMM can effectively 
discover a variety of interesting interpretable student behavior patterns at different levels of resolution, many of which are beyond what existing approaches can discover. 2L-HMM further enables easy use of the discovered patterns for discriminative analysis such as prediction of learning outcome. 
Since our proposed methods are unsupervised, they can potentially be applied to any MOOC data without requiring any manual work to facilitate understanding of student behaviors and their variations, opening up many possibilities for developing intelligent tutoring systems that can adaptive to student behavior.

### 2017-04-04 sec-model.tex
Switch to 2L-HMM for the model abbreviation.
Update the reviewer response draft to mention this concrete change.

Adding this second layer would allow us to characterize a latent task
state in much more detail, as it would reveal not only what activities are
most common to a task state, but also the transition patterns between these
``micro'' activities (e.g., it can reveal frequent back-and-forth
transitions between quiz-taking and quiz-submission, which would suggest a
concentrated period of taking quizzes). Combining this ``surface'' Markov
model over the ``micro'' actions with the ``deep'' hidden Markov model over
the latent task states gives us a fairly general and powerful two-layer HMM
(TL-HMM) that can simultaneously learn ``deeply'' the latent task/behavior
states and their transitions as well as the corresponding ``micro''
activity transition patterns associated with each latent state to
facilitate interpretation and analysis of the discovered latent state
patterns. Next, we present this model more formally and discuss how to
estimate its parameters to uncover these latent patterns in an unsupervised
manner.

\subsection{Formal Definition of the TL-HMM}

@@

Fortunately, the recursions for the $\alpha$ and $\beta$ variables remain
the same, as do the definitions of $\gamma_t(i)$ and $\xi_t(i,j)$ in the
E-step. The only change is in the updating equations in the M-step,
where we replace the update for $b_i(v)$ by a pair of updates: one for
$\pi^{(i)}_a$ one for $A^{(i)}_{ab}$. We denote the $m$-th action in the
$t$-th sequence for the $k$-th student as $o_{k,t,m}$. The two updates can
be written as
\begin{align}
  \pi^{(i)}_{a}
  &= \frac{\sum_{k=1}^M \sum_{t=1,a_1 = a}^T \gamma^{(\mathbf{o}_k)}_t(i)}
  {\sum_{k=1}^M \sum_{t=1}^T \gamma^{(\mathbf{o}_k)}_t(i)}, \text{ and }\\
  A^{(i)}_{ab}
  &= \frac{\sum_{k=1}^M \sum_{t=1}^T
  \sum_{m=2,a_{m-1}=a \land a_m = b}^{|\mathbf{o}_{k,t}|}
  \gamma^{(\mathbf{o}_k)}_t(i)}
  {\sum_{k=1}^M \sum_{t=1}^T \sum_{m=2}^{|\mathbf{o}_{k,t}|}
  \gamma^{(\mathbf{o}_k)}_t(i)}
\end{align}
where $a_m = o_{t,k,m}$ (for notational conciseness). Our modified EM
algorithm for TL-HMMs is provided as part of the \meta/
toolkit~\citemeta{}.

>>>>

Adding this second layer would allow us to characterize a latent task
state in much more detail, as it would reveal not only what activities are
most common to a task state, but also the transition patterns between these
``micro'' activities (e.g., it can reveal frequent back-and-forth
transitions between quiz-taking and quiz-submission, which would suggest a
concentrated period of taking quizzes). Combining this ``surface'' Markov
model over the ``micro'' actions with the ``deep'' hidden Markov model over
the latent task states gives us a fairly general and powerful two-layer HMM
(2L-HMM) that can simultaneously learn ``deeply'' the latent task/behavior
states and their transitions as well as the corresponding ``micro''
activity transition patterns associated with each latent state to
facilitate interpretation and analysis of the discovered latent state
patterns. Next, we present this model more formally and discuss how to
estimate its parameters to uncover these latent patterns in an unsupervised
manner.

\subsection{Formal Definition of the 2L-HMM}

@@

Fortunately, the recursions for the $\alpha$ and $\beta$ variables remain
the same, as do the definitions of $\gamma_t(i)$ and $\xi_t(i,j)$ in the
E-step. The only change is in the updating equations in the M-step,
where we replace the update for $b_i(v)$ by a pair of updates: one for
$\pi^{(i)}_a$ one for $A^{(i)}_{ab}$. We denote the $m$-th action in the
$t$-th sequence for the $k$-th student as $o_{k,t,m}$. The two updates can
be written as
\begin{align}
  \pi^{(i)}_{a}
  &= \frac{\sum_{k=1}^M \sum_{t=1,a_1 = a}^T \gamma^{(\mathbf{o}_k)}_t(i)}
  {\sum_{k=1}^M \sum_{t=1}^T \gamma^{(\mathbf{o}_k)}_t(i)}, \text{ and }\\
  A^{(i)}_{ab}
  &= \frac{\sum_{k=1}^M \sum_{t=1}^T
  \sum_{m=2,a_{m-1}=a \land a_m = b}^{|\mathbf{o}_{k,t}|}
  \gamma^{(\mathbf{o}_k)}_t(i)}
  {\sum_{k=1}^M \sum_{t=1}^T \sum_{m=2}^{|\mathbf{o}_{k,t}|}
  \gamma^{(\mathbf{o}_k)}_t(i)}
\end{align}
where $a_m = o_{t,k,m}$ (for notational conciseness). Our modified EM
algorithm for 2L-HMMs is provided as part of the \meta/
toolkit~\citemeta{}.

### 2017-04-04 sec-results.tex
Fix a few compile errors.
Not sure what changed here; I was able to compile it before...
Move table captions to occur on top of the tables.
JEDM style guide apparently specifies this.
Switch to 2L-HMM for the model abbreviation.
Update the reviewer response draft to mention this concrete change.

\section{Experiment Results}
As an analysis tool, the TL-HMM model provides with us the following two
patterns to characterize student behavior:
\begin{enumerate*}[label=(\arabic*)]
  \item the latent state representations, and
  \item the latent state transitions.
\end{enumerate*}
Thus to evaluate the proposed model, we conduct experiments to
qualitatively analyze both types of patterns discovered from empirical MOOC
log data.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{rrrr}
      \textbf{MOOC} & \textbf{Students} & \textbf{Sequences} & \textbf{Avg.
      $|\mathbf{s}|$}\\\hline
      \textretrieval{} & 18,941 & 85,240 & 7.31\\
      \sustain{} & 85,240 & 231,881 & 15.4
    \end{tabular}
    \caption{Statistics about the sequences extracted from the two MOOCs.}
    \label{table:datasets}
  \end{center}
\end{table}

\subsection{Latent State Representations}
The TL-HMM is meant to be a tool for exploratory analysis of student
behavior. As such, the number of states should be empirically set based on
the goal of analysis. A higher number of states will lead to a
finer-grained modeling of student behavior. In our experiments, we explored
using between 2--6 states. First, we fit a 6-state TL-HMM to the
\textretrieval{} sequence dataset and show some of the latent state
representations we find. We used the following ten actions as our action
set $\mathbf{A}$:
\begin{enumerate*}[label=(\arabic*)]
  \item quiz start,
  \item quiz submit,
  \item wiki (course material),
  \item forum list (view the list of all forums),
  \item forum thread list (view the list of all threads in a specific
    forum),
  \item forum thread view (view the list of posts within a specific
    thread),
  \item forum search (a search query issued against the forum),
  \item forum post thread (a new thread was posted),
  \item forum post reply (a new post was created within an existing
    thread), and
  \item view lecture (defined as either streaming or downloading a lecture
    video).
\end{enumerate*}

@@

\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/text-6state/state0.png}
    \caption{An example ``quiz taking'' state.}
    \label{fig:practice-quiz-state}
  \end{subfigure}%
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/text-6state/state1.png}
    \caption{An example ``lecture viewing'' state.}
    \label{fig:lecture-viewing-state}
  \end{subfigure}
  \caption{Two example states found by the 6-state TL-HMM. (The naming of
  these states reflects our own interpretation.)}
  \label{fig:states}
\end{figure*}

\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/text-6state/state4.png}
    \caption{A state from \textretrieval{}.}
  \end{subfigure}%
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/sustain-6state/state4.png}
    \caption{A state from \sustain{}.}
  \end{subfigure}
  \caption{Two similar example states found by the 6-state TL-HMM.}
  \label{fig:cross-course}
\end{figure*}

@@

We can also use our model for cross-course behavior analysis. In
Figure~\ref{fig:cross-course} we show two latent state representations
learned by a 6-state TL-HMM, one from \textretrieval{} and one from
\sustain{}. These two states were chosen as they are the most similar
across the two courses. However, if we look at the transitions we can see
some important differences. First, in the state from \textretrieval{}, we
can see that the probability of returning to the course wiki after viewing
a lecture (transition probability $p \approx 0.24$) is considerably lower
than that probability in the state from \sustain{} (transition probability
$p \approx 0.57$). We also notice that the self-loop for staying in a
lecture activity in \textretrieval{} (transition probability $p \approx
0.70$) is significantly higher probability than it is in the state from
\sustain{} (transition probability $p \approx 0.34$).  This gives us some
insight into the lecture viewing behavior of students in these two MOOCs
which can possibly be a reflection of the course's structure (which, as
demonstrated in \citet{Davis:2016:EDM}, can differ substantially across
different MOOCs).  In \textretrieval{}, students are likely to view lecture
videos in succession directly without first visiting the course wiki,
whereas in \sustain{} students are much more likely to first return the
course wiki before watching the next lecture video. This observation would
be lost if we did not consider the transitions between the actions within
the latent states.

\subsection{Varying the Number of Latent States}
Our TL-HMM model has a parameter $K$ that sets the number of latent states
to be learned. We believe that this can allow our model to flexibly
discover patterns of different granularity, and we can show this by varying
$K$ for a particular course and observing how the latent state
representations evolve.

@@

We observe similar behavior in Figure~\ref{fig:sustain-state-evolution}
as we increase $K$ when fitting our TL-HMM on the \sustain{} MOOC.\ Again,
in the transition between $K=2$ and $K=3$ we discover forum behavior
patterns as a latent state. However, in the transition between $K=3$ to
$K=4$ we instead see a refining of that discovered forum state, where
state 3 captures asymmetric transition probabilities between ``forum thread
view'' and ``forum thread list''. These states could be seen as redundant,
in which case a  setting of $K=3$ may be more appropriate for the
\sustain{} dataset.

@@

A unique property of our model is its ability to capture transitions
between the \emph{behavior patterns themselves} that are captured by the
latent states. In Figure~\ref{fig:trans-avg} we show the latent state
transition diagram for a 3-state TL-HMM fit on \textretrieval{}. We can
immediately observe two things: \begin{enumerate*}[label=(\arabic*)]
  \item each latent state has a very high ``staying'' probability, and
  \item the prevalence of each latent state matches our intuition.
\end{enumerate*}
In particular, we can see that the forum browsing state (state 2) has
relatively lower probability than the other states as we might expect. It
also makes sense that state 0 has rather high probability as this state
likely captures all sequences where a student logged in to the platform and
then did nothing else (likely checking for updates). If we look at state 1
and state 3, their relative probabilities match our intuition as
well. State 1 seems to capture a more engaged browsing session, where there
is non-negligible probability associated with different activities such as
quiz taking and forum browsing and, importantly, these activities have high
probability symmetric edges (so students are taking quizzes one after the
other, or viewing forum threads in succession). By contrast, state 3 seems
to capture a more passive student, with negligible probability mass
associated with forum activity (with low symmetry in the edges). The link
between ``quiz submit'' and ``quiz start'' (indicating quiz repetition) is
also significantly lower than state 1.

\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth,trim={0 3cm 0 2cm}]{figures/trans-comp/trans-avg.png}
    \caption{\label{fig:trans-avg}}
  \end{subfigure}%
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth,trim={0, 3cm 0 2cm}]{figures/trans-comp/trans-perfect.png}
    \caption{\label{fig:trans-perfect}}
  \end{subfigure}%
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth,trim={0, 3cm 0 2cm}]{figures/trans-comp/trans-low.png}
    \caption{\label{fig:trans-low}}
  \end{subfigure}
  \caption{The latent state transition diagrams for a 4-state TL-HMM fit to
  \protect\textretrieval{} for all students (a) compared to only ``perfect''
  students (b) and only ``low'' students (c).}
  \label{fig:trans-comp}
\end{figure*}

Thus, we might expect to see students that perform well in the course
preferring states 1 and 2 over states 0 and 3. To verify this, we took the
model we learned on the full training data and retrofit it to training data
consisting only of sequences produced by students in \textretrieval{} that
had perfect marks. To prevent the latent state meanings from drifting, we
forced the model parameters associated with their Markov model
representations to be fixed, in effect only learning initial and transition
probabilities for the top layer of our TL-HMM.\ We show the updated latent
state transition diagram in Figure~\ref{fig:trans-perfect}. We can clearly
see that the probability of state 2 has increased dramatically, consistent
with previous observations of the positive correlation between forum activity and
grades~\cite{Huang:2014:LAS}, while the probability of states 0 and 3 has
decreased. State 1 had its probability increase, but only very slightly.

@@

\begin{table}
  \centering
  \begin{tabular}{r|rrrrr}
    \textbf{Group} & \textbf{State 0} & \textbf{State 1}^\dagger &
    \textbf{State 2}^\dagger & \textbf{State 3} & \textbf{State 2
    $\rightarrow$ 2}^\dagger\\\hline
    \textbf{Perfect} & 975.3  & 1001.5 & 999.0  & 1056.5 & 939.6\\
    \textbf{Low}     & 1024.9 & 816.4  & 1230.5 & 1161.2 & 1187.4\\
  \end{tabular}
  \caption{Average rank for ``perfect'' and ``low'' student groups in the
  ranked lists associated with the four latent states found by a TL-HMM.
  $\dagger$ indicates statistically significant different mean ranks at $p
  < 0.01$ according to an unpaired $t$-test.}
  \label{table:mean-rank}
\end{table}

>>>>

\section{Experiment Results}
As an analysis tool, the 2L-HMM model provides with us the following two
patterns to characterize student behavior:
\begin{enumerate*}[label=(\arabic*)]
  \item the latent state representations, and
  \item the latent state transitions.
\end{enumerate*}
Thus to evaluate the proposed model, we conduct experiments to
qualitatively analyze both types of patterns discovered from empirical MOOC
log data.

@@

\begin{table}
  \begin{center}
    \caption{Statistics about the sequences extracted from the two MOOCs.}
    \label{table:datasets}
    \begin{tabular}{rrrr}
      \textbf{MOOC} & \textbf{Students} & \textbf{Sequences} & \textbf{Avg.
      $|\mathbf{s}|$}\\\hline
      \textretrieval{} & 18,941 & 85,240 & 7.31\\
      \sustain{} & 85,240 & 231,881 & 15.4
    \end{tabular}
  \end{center}
\end{table}

\subsection{Latent State Representations}
The 2L-HMM is meant to be a tool for exploratory analysis of student
behavior. As such, the number of states should be empirically set based on
the goal of analysis. A higher number of states will lead to a
finer-grained modeling of student behavior. In our experiments, we explored
using between 2--6 states. First, we fit a 6-state 2L-HMM to the
\textretrieval{} sequence dataset and show some of the latent state
representations we find. We used the following ten actions as our action
set $\mathbf{A}$:
\begin{enumerate*}[label=(\arabic*)]
  \item quiz start,
  \item quiz submit,
  \item wiki (course material),
  \item forum list (view the list of all forums),
  \item forum thread list (view the list of all threads in a specific
    forum),
  \item forum thread view (view the list of posts within a specific
    thread),
  \item forum search (a search query issued against the forum),
  \item forum post thread (a new thread was posted),
  \item forum post reply (a new post was created within an existing
    thread), and
  \item view lecture (defined as either streaming or downloading a lecture
    video).
\end{enumerate*}

@@

\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/text-6state/state0.png}
    \caption{An example ``quiz taking'' state.}
    \label{fig:practice-quiz-state}
  \end{subfigure}%
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/text-6state/state1.png}
    \caption{An example ``lecture viewing'' state.}
    \label{fig:lecture-viewing-state}
  \end{subfigure}
  \caption{Two example states found by the 6-state 2L-HMM. (The naming of
  these states reflects our own interpretation.)}
  \label{fig:states}
\end{figure*}

\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/text-6state/state4.png}
    \caption{A state from \textretrieval{}.}
  \end{subfigure}%
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/sustain-6state/state4.png}
    \caption{A state from \sustain{}.}
  \end{subfigure}
  \caption{Two similar example states found by the 6-state 2L-HMM.}
  \label{fig:cross-course}
\end{figure*}

@@

We can also use our model for cross-course behavior analysis. In
Figure~\ref{fig:cross-course} we show two latent state representations
learned by a 6-state 2L-HMM, one from \textretrieval{} and one from
\sustain{}. These two states were chosen as they are the most similar
across the two courses. However, if we look at the transitions we can see
some important differences. First, in the state from \textretrieval{}, we
can see that the probability of returning to the course wiki after viewing
a lecture (transition probability $p \approx 0.24$) is considerably lower
than that probability in the state from \sustain{} (transition probability
$p \approx 0.57$). We also notice that the self-loop for staying in a
lecture activity in \textretrieval{} (transition probability $p \approx
0.70$) is significantly higher probability than it is in the state from
\sustain{} (transition probability $p \approx 0.34$).  This gives us some
insight into the lecture viewing behavior of students in these two MOOCs
which can possibly be a reflection of the course's structure (which, as
demonstrated in \citet{Davis:2016:EDM}, can differ substantially across
different MOOCs).  In \textretrieval{}, students are likely to view lecture
videos in succession directly without first visiting the course wiki,
whereas in \sustain{} students are much more likely to first return the
course wiki before watching the next lecture video. This observation would
be lost if we did not consider the transitions between the actions within
the latent states.

\subsection{Varying the Number of Latent States}
Our 2L-HMM model has a parameter $K$ that sets the number of latent states
to be learned. We believe that this can allow our model to flexibly
discover patterns of different granularity, and we can show this by varying
$K$ for a particular course and observing how the latent state
representations evolve.

@@

We observe similar behavior in Figure~\ref{fig:sustain-state-evolution}
as we increase $K$ when fitting our 2L-HMM on the \sustain{} MOOC.\ Again,
in the transition between $K=2$ and $K=3$ we discover forum behavior
patterns as a latent state. However, in the transition between $K=3$ to
$K=4$ we instead see a refining of that discovered forum state, where
state 3 captures asymmetric transition probabilities between ``forum thread
view'' and ``forum thread list''. These states could be seen as redundant,
in which case a  setting of $K=3$ may be more appropriate for the
\sustain{} dataset.

@@

A unique property of our model is its ability to capture transitions
between the \emph{behavior patterns themselves} that are captured by the
latent states. In Figure~\ref{fig:trans-avg} we show the latent state
transition diagram for a 3-state 2L-HMM fit on \textretrieval{}. We can
immediately observe two things: \begin{enumerate*}[label=(\arabic*)]
  \item each latent state has a very high ``staying'' probability, and
  \item the prevalence of each latent state matches our intuition.
\end{enumerate*}
In particular, we can see that the forum browsing state (state 2) has
relatively lower probability than the other states as we might expect. It
also makes sense that state 0 has rather high probability as this state
likely captures all sequences where a student logged in to the platform and
then did nothing else (likely checking for updates). If we look at state 1
and state 3, their relative probabilities match our intuition as
well. State 1 seems to capture a more engaged browsing session, where there
is non-negligible probability associated with different activities such as
quiz taking and forum browsing and, importantly, these activities have high
probability symmetric edges (so students are taking quizzes one after the
other, or viewing forum threads in succession). By contrast, state 3 seems
to capture a more passive student, with negligible probability mass
associated with forum activity (with low symmetry in the edges). The link
between ``quiz submit'' and ``quiz start'' (indicating quiz repetition) is
also significantly lower than state 1.

\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth,trim={0 3cm 0 2cm}]{figures/trans-comp/trans-avg.png}
    \caption{\label{fig:trans-avg}}
  \end{subfigure}%
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth,trim={0, 3cm 0 2cm}]{figures/trans-comp/trans-perfect.png}
    \caption{\label{fig:trans-perfect}}
  \end{subfigure}%
  \begin{subfigure}[t]{0.33\textwidth}
    \includegraphics[width=\textwidth,trim={0, 3cm 0 2cm}]{figures/trans-comp/trans-low.png}
    \caption{\label{fig:trans-low}}
  \end{subfigure}
  \caption{The latent state transition diagrams for a 4-state 2L-HMM fit to
  \protect\textretrieval{} for all students (a) compared to only ``perfect''
  students (b) and only ``low'' students (c).}
  \label{fig:trans-comp}
\end{figure*}

Thus, we might expect to see students that perform well in the course
preferring states 1 and 2 over states 0 and 3. To verify this, we took the
model we learned on the full training data and retrofit it to training data
consisting only of sequences produced by students in \textretrieval{} that
had perfect marks. To prevent the latent state meanings from drifting, we
forced the model parameters associated with their Markov model
representations to be fixed, in effect only learning initial and transition
probabilities for the top layer of our 2L-HMM.\ We show the updated latent
state transition diagram in Figure~\ref{fig:trans-perfect}. We can clearly
see that the probability of state 2 has increased dramatically, consistent
with previous observations of the positive correlation between forum activity and
grades~\cite{Huang:2014:LAS}, while the probability of states 0 and 3 has
decreased. State 1 had its probability increase, but only very slightly.

@@

\begin{table}
  \centering
  \caption{Average rank for ``perfect'' and ``low'' student groups in the
  ranked lists associated with the four latent states found by a 2L-HMM.
  $\dagger$ indicates statistically significant different mean ranks at $p
  < 0.01$ according to an unpaired $t$-test.}
  \label{table:mean-rank}
  \begin{tabular}{r|rrrrr}
    \textbf{Group} & \textbf{State 0} & \textbf{State
  1}\textsuperscript\textdagger &
    \textbf{State 2}\textsuperscript\textdagger & \textbf{State 3} & \textbf{State 2
    $\rightarrow$ 2}\textsuperscript\textdagger\\\hline
    \textbf{Perfect} & 975.3  & 1001.5 & 999.0  & 1056.5 & 939.6\\
    \textbf{Low}     & 1024.9 & 816.4  & 1230.5 & 1161.2 & 1187.4\\
  \end{tabular}
\end{table}
